Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-02-23 alloscore_manuscript/alloscore-application-shorter.Rnw
edits to text

Decision theory investigates how decisions are made by formalizing a decision as the process of selecting an
\emph{action} $x$ from a specified set $\mathcal{X}$ of potential actions while taking into consideration the possible
consequences of $x$ under future states of the world. Let $Y$ be a quantifiable aspect of the future which will help to
determine the consequences of $x$. $Y$ is a random variable inasmuch as it is indeterminate when the decision is made.
We refer to a realization $y$ of $Y$ as an \emph{outcome}. When a decision maker assumes $Y$ to have a specified
distribution $F_Y$ and $Y$ is the only aspect of the future relevant to the selection of $x$, the process is referred to
as a decision under risk.  The selection of $x$ without a full specification of the distribution of $Y$ is referred to
as a decision under uncertainty. Clearly almost any real-world decision will be a decision under uncertainty.  Yet many
foundational results in decision theory were guided (as we will be) by the intuition that much can be learned about a
decision making process by building a ``small world'' approximation of the real-world (after \cite{savage1954foundations})
in which the beliefs of a decision
maker are equivalent to a fully specified distribution of a $Y$ that exhaustiveley quantifies all relevant aspects of
the future, converting a decision under uncertainty to a decison under risk.

Examples of actions with which this paper is concerned are levels of investment in measures designed to mitigate severe
disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of
all feasible levels of investment. An outcome $y$ determining the consequences of such an action might be the number of
individuals who eventually become sick and would benefit from the mitigation measure. A fundamental strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numerical value, or
\emph{loss}, measuring the (lack of) success of $x$. In our example, $x$ will succeed to the extent that it meets the
realized need $y$, so that greater unmet need will incur greater loss.

One tool a decision maker can use to anticipate the consequences of an action $x$ is a forecast
$F$ of $Y$. And within a ``small world'' where $Y$ exhaustively quantifies the future, the decision maker might choose that
$x$ which yields the lowest expected loss according to $F$. The loss of the action $x$ chosen according to $F$ can then, 
in turn, be interpreted as the value of $F$ as an
input to the decision making process under the realized outcome $y$ of $Y$.

@@

Probabilistic forecasts of infectious disease outbreaks have typically been evaluated using well-known proper scoring
rules such as the LS, CRPS, and WIS. Often models are ranked according to such a score, but without reference to any
underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated.
We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by
refocussing the attention of modelers and public health officials involved in these efforts on the elicitation of
forecasts designed to support specific public health decision problems, and moreover, the evaluation of forecasts with
scoring rules that directly consider the success of forecasts in supporting these decisions. As evidence for this
possibility, we have demonstrated how tying forecast evaluation to a specific decision problem (via the allocation
score) can yield model rankings that differ substantively from those based a current standard measure of forecast
accuracy, the WIS, which does not refer to any decision problem that has been clearly connected to outbreak response.
This result aligns with findings in other fields, especially those where the monetary consequences of using a forecast
can be clearly juxtaposed with standard accuracy measures of the forecast \citep{leitch1991economicForecastEval,
murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}. Additionally, we show that an existing
ensemble forecast approach was the only method to outperform (in terms of this new evaluation method) two benchmark
allocation approaches in a hypothetical application, suggesting that there is room for innovation of current
epidemiological modeling techniques that might be spurred on by a reorientation of the field toward practical decision
problems. For example, in order to surpass benchmark allocation scores, forecasters need to devote effort to capturing
the \emph{relative} magnitude of future resource need across different locations.  Success at this task is not, however,
directly rewarded by the WIS.

>>>>

Decision theory investigates how decisions are made by formalizing a decision as the process of selecting an
\emph{action} $x$ from a specified set $\mathcal{X}$ of potential actions while taking into consideration the possible
consequences of $x$ under future states of the world. Let $Y$ be a quantifiable aspect of the future which will help to
determine the consequences of $x$. $Y$ is a random variable inasmuch as it is indeterminate when the decision is made.
We refer to a realization $y$ of $Y$ as an \emph{outcome}. A fundamental strategy in decision
theory is to assume that the consequences of an action $x$ under outcome $y$ can be assigned a numeric value, or
\emph{loss}, measuring the (lack of) success of $x$.

Some examples of actions are levels of investment in measures designed to mitigate severe
disease outcomes such as hospital beds, ventilators, medication, or medical staff, with $\mathcal{X}$ being the set of
all feasible levels of investment. An outcome $y$ determining the consequences of such an action might be the number of
individuals who eventually become sick and would benefit from the mitigation measure. In our example, $x$ will succeed to
the extent that it meets the realized need $y$, so that greater unmet need will incur greater loss.

When a forecast $F$ of $Y$ is available, a decision maker might choose that action
$x$ which yields the lowest expected loss according to $F$.
The loss of the action $x$ chosen according to $F$ can then, in turn, be interpreted as the value of $F$ as an
input to the decision making process under the realized outcome $y$ of $Y$.

The strategy of minimizing the expected loss according to $F$ might be used if the
decision maker trusted that $F$ was accurate and did not have any additional information that was not captured
in $F$. However, here we use it only as a device for evaluating forecasts by examining the consequences of
hypothetical decisions that would be made when using only $F$ as an input.
We return to this point in the discussion.

@@

Probabilistic forecasts of infectious disease outbreaks have typically been evaluated using well-known proper scoring
rules such as the LS, CRPS, and WIS. Often models are ranked according to such a score, but without reference to any
underlying decision problem from which the score might be derivable as a Bayes scoring rule or be otherwise motivated.
We argue that the value of collaborative forecasting efforts, such as the COVID-19 Forecasting Hub, could be enhanced by
including evaluations of forecasts with
scoring rules that directly consider the success of forecasts in supporting specific public health decisions. As evidence for this
possibility, we have demonstrated how tying forecast evaluation to a specific decision problem (via the allocation
score) can yield model rankings that differ substantively from those based a current standard measure of forecast
accuracy, the WIS, which does not refer to any decision problem that has been clearly connected to outbreak response.
This result aligns with findings in other fields, especially those where the monetary consequences of using a forecast
can be clearly juxtaposed with standard accuracy measures of the forecast \citep{leitch1991economicForecastEval,
murphy1993whatisagoodforecast, cenesizoglu2012returnPredictionEconValue}. Additionally, we show that an existing
ensemble forecast approach was the only method to outperform (in terms of this new evaluation method) two benchmark
allocation approaches in a hypothetical application, suggesting that there is room for innovation of current
epidemiological modeling techniques that might be spurred on by a reorientation of the field toward practical decision
problems. For example, in order to surpass benchmark allocation scores, forecasters need to devote effort to capturing
the \emph{relative} magnitude of future resource need across different locations.  Success at this task is not, however,
directly rewarded by the WIS.
